HPC--101\\Onboarding

### 2016-01-21 hpc_101/paper.tex
Added slide simulating submission of example script removed some commented text, and re worked the Job scheduler definition.

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module unload $<$module name$>$' -- Remove named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
		%\begin{itemize}\footnotesize
			%\item Two ways to install software on the cluster
			%\begin{itemize}\tiny
				%\item Download compatible binaries
				%\item Compile software from source
			%\end{itemize}
		%\end{itemize}
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

@@

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	\begin{definition}{\footnotesize
		A job scheduler is a computer application for controlling unattended background program execution (commonly called batch processing)
		}
	\end{definition}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize

        \item The ability for the user to interact with the schduler to define workflows and/or define job dependencies
        \item[--] Interfaces which help to define workflows and/or job dependencies

        \item The ability to automatically manage the scheduling of user jobs
        \item[--] Automatic submission of executions

        \item The functionality to allow users to monitor submitted jobs
        \item[--] Interfaces to monitor the executions

        \item The capability to control and prioritize the execution order of unrelated jobs
        \item[--] Priorities and/or queues to control the execution order of unrelated jobs
	\end{itemize}
	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.  Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job   
\end{block}
%The following slides that cover slurm will cover
%\begin{enumerate}
%\item Several useful SLURM commands
%\item Resource partitions
%\item example interactive jobs and batch job scripts
%\end{enumerate}
\end{frame}

@@

\frametitle{SLURM Job Scripts}
\begin{frame}
\frametitle{Interactive Job with SLURM}
	\begin{block}{Interactive session (no X11)}\tiny
	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00 \ddash{}pty /bin/bash
	\end{block}

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

@@

\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm

@@

>>>>

	\end{block}
	\begin{itemize}
		\item We globally install frequently requested software packages and create modules for all users to access
		\item Access to modules is via the \textbf{module} command
		\begin{itemize}\footnotesize
			\item `module avail' -- list installed modules
			\item `module show $<$module name$>$' -- Show what actions a module performs
			\item `module load $<$module name$>$' -- Loads the named module
			\item `module purge' -- Unload all loaded modules
		\end{itemize}
		\item Installing software in your \ctilde{}/apps directory is suggested to prevent us from being a bottleneck
	\end{itemize}
	\footnotetext[1]{\tiny \url{https://en.wikipedia.org/wiki/Environment_Modules_(software)}}
\end{frame}

@@

\begin{frame}
	\frametitle{What is a Job Scheduler?}
	Basic features expected of a job scheduler include:
	\begin{itemize}\footnotesize
        \item The capability to control and prioritize the execution order of unrelated jobs
        \item Allocates exclusive and/or non-exclusive access to resources to users for some duration of time so they can perform work
				\item Provides a framework for starting, executing, and monitoring work on the set of allocated nodes
				\item Arbitrate contention for resources by managing a queue of pending work
	\end{itemize}
	\bigskip
	For the {\craycs}, we utilize the \textbf{S}imple \textbf{L}inux \textbf{U}tility \textbf{R}esource \textbf{M}anager~\\ or simply known as the \emph{SLURM scheduler}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Job_scheduler}}
	\numlessfootnotetxt{\tiny \url{https://en.wikipedia.org/wiki/Slurm_Workload_Manager}}
	\numlessfootnotetxt{\tiny \url{http://slurm.schedmd.com/slurm.html}}
\end{frame}

\subsection{SLURM}
\begin{frame}
\frametitle{SLURM}
\begin{block}{How are jobs scheduled?}
User submitted jobs are assigned a priority using a fairshare algorithm.  Factors such as: runtime, resource usage request, age of job, and amount of compute time a user has used in recent history are all factored in to assigning a priority to a give job   
\end{block}
\end{frame}

@@

\subsection{Interactive Job -- srun}
\frametitle{SLURM Job Scripts}
\begin{frame}
\frametitle{Interactive Job with SLURM}
	\begin{block}{Interactive session (no X11)}\tiny
	$[$login \ctilde$]$\$ srun \ddash{}immediate \ddash{}partition sb.q \ddash{}nodes 1 \ddash{}cpus-per-task 5 \ddash{}tasks-per-node 1 \ddash{}time 0-01:00 \ddash{}pty /bin/bash
	\end{block}

@@

\subsection{MPI Batch Job Example Script}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat mpi.slurm

@@

\subsection{Non-MPI Batch Job Example Script}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Non-MPI Job)}
\begin{semiverbatim}\tiny
[login lus]\$ cat hello_world.slurm

@@

\subsection{Batch Job Example Script Using Job Array}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Submission Script File (Job Array)}
\begin{semiverbatim}\tiny
[login lus]\$ cat job_array.slurm

@@

\subsection{Executing and Monitoring Batch Jobs}
\begin{frame}[fragile]
\frametitle{SLURM sbatch -- Executing \& Monitoring Jobs}\footnotesize
To execute a submission script you use the \textbf{\texttt{sbatch}} command
\begin{block}{Example -- Non Job Array}
\begin{semiverbatim}\tiny
[login lus]\$ sinfo
PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
[login lus]\$ sbatch hello_world.slurm
sbatch: Submitted batch job 469
[login lus]\$ squeue
JOBID PARTITION    NAME     USER    ST TIME  NODES  NODELIST(REASON)
469   community.q  example  user99  R  00:01 1      compute-0001
\end{semiverbatim}
\end{block}
\begin{block}{Example -- Job Array}
\begin{semiverbatim}\tiny
[login lus]\$ sinfo
PARTITION     AVAIL  TIMELIMIT  NODES  STATE NODELIST
community.q   up     3-00:00:00 2      idle  compute-[0001-0002]
[login lus]\$ sbatch {\ddash}array=0-2 job_array.slurm
sbatch: Submitted batch job 470
[login lus]\$ squeue
JOBID     PARTITION    NAME     USER    ST   TIME    NODES  NODELIST(REASON)
470[0-2]  community.q  example  user99  PD   00:00   1      (resource)
470_0     community.q  example  user99  R    00:01   1      compute-0001
470_1     community.q  example  user99  R    00:05   1      compute-0002
\end{semiverbatim}
\end{block}
